**SV2414 and vrp2113**
**Energy Storage Options in Sao Vicente**
Note: This RMD contains setpoint calculation, along with question 5 of 'Storage Capacity Effects on Capacity Factor' section (which uses setpoints).
```{r, echo=FALSE}
setwd("C:/Users/Sagar/OneDrive/Documents/EI/Final")
if(!require(plyr)) install.packages("plyr")
library(plyr)
if(!require(ggplot2)) install.packages("ggplot2")
library(ggplot2)

#Import data
dem<-read.csv("sao.vicente.load.csv")
wind<-read.csv("sao.vicente.wind.csv")
mpc<-read.csv("v52-850KW-power-curve.csv")

#Rename columns
names(wind)[6]<-"del_hourly"
names(wind)[5]<-"windspeed"
names(wind)[4]<-"windgen"

#Interpolate for MPC power 
interp=splinefun(mpc$windspeed_mps, mpc$power_kW,method = "natural")
#Find equivalent MPC power for the windspeeds we have
wind$mpc_power<-interp(wind$windspeed)*7

#Reassign values that are greater than MPC power equivalents to MPC power equivalents
for (j in 1:nrow(wind)){
  if(wind$windgen[j]>wind$mpc_power[j])
    wind$windgen[j]<-wind$mpc_power[j]
}

#Convert date to date format
wind$date<-as.Date(wind$date,format="%Y-%m-%d")
#Create new column for datetime using date and hour
wind$datetime=as.POSIXct(wind$date+wind$hour*(1/24)+(1/3), format="%Y-%m-%d %H:%M:%S", origin="2013-01-01 00:00:00")

#Convert date and time to datetime format for demand data
dem$data.pt2.date.time<-as.POSIXct(dem$data.pt2.date.time,format="%Y-%m-%d %H:%M:%S")
#Rename columns
names(dem)[3]<-"datetime"
names(dem)[4]<-"demand"

#Remove columns we don'tneed 
wind<-wind[-c(1,2,3,7)]
dem<-dem[-c(1,2)]

#Merge generation and demand data
ad<-merge(x = dem,y = wind,by="datetime",all.x=TRUE)

#Fill missing generation data
ad$windspeed[is.na(ad$windspeed)]<-summary(wind$windspeed)[4]
ad$del_hourly[is.na(ad$del_hourly)]<-0
for (i in 1:nrow(ad)){
  if(is.na(ad$windgen[i])){
    ad$windgen[i]<-interp(wind$windspeed[i])*7
    }
}

#Convert negative generation values to 0
ad$del_hourly[ad$del_hourly<0]<-0
ad$windgen[ad$windgen<0]<-0

#Correct small demand values to mean demand
ad$demand[ad$demand<quantile(ad$demand,0.005)]<-summary(ad$demand)[4]
```
###**REFINING SYSTEM SETPOINTS**
**1. Propose a reasonable set of setpoints for wind farm generation and baseload generation. How do these values affect the curtailment and capacity factor?**
```{r}
#Average demand in the evenings
avgeved=summary(ad$demand[(as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour>=17 & (as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour<=21])[4]
#Average demand in the day
avgdayd=summary(ad$demand[(as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour>=6 & (as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour<=16])[4]
#Average demand at late nights
avglnd=summary(ad$demand[(as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour>=22 | (as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour<=05])[4]

#Average windgen in the evenings
avgevegen=summary(ad$windgen[(as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour>=17 & (as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour<=21])[4]
#Average windgen in the day
avgdaygen=summary(ad$windgen[(as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour>=6 & (as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour<=16])[4]
#Average windgen at late nights
avglngen=summary(ad$windgen[(as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour>=22 | (as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour<=05])[4]

#Average deficit in the evenings, day and late nights
avgevedef=avgeved-avgevegen
avgdaydef=avgdayd-avgdaygen
avglndef=avglnd-avglngen

#Assign setpoints for baseload
for (i in 1:nrow(ad)){
  if((as.POSIXlt(ad$datetime[i],format="%Y-%m-%d %H:%M:%S"))$hour>=17 & (as.POSIXlt(ad$datetime[i],format="%Y-%m-%d %H:%M:%S"))$hour<=21){
    ad$blsp[i]=avgevedef
    }
  if((as.POSIXlt(ad$datetime[i],format="%Y-%m-%d %H:%M:%S"))$hour>=6 & (as.POSIXlt(ad$datetime[i],format="%Y-%m-%d %H:%M:%S"))$hour<=16){
    ad$blsp[i]=avgdaydef
    }
  if((as.POSIXlt(ad$datetime[i],format="%Y-%m-%d %H:%M:%S"))$hour>=22 | (as.POSIXlt(ad$datetime[i],format="%Y-%m-%d %H:%M:%S"))$hour<=05){
    ad$blsp[i]=avglndef
  }
  #Find gap to be filled above baseload
  ad$gap[i]=max(ad$demand[i]-ad$blsp[i],0.0)
}

#Find average gap to be filled in evenings, days and late nights, and take appropriate fraction
avgevegap=0.76*summary(ad$gap[(as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour>=17 & (as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour<=21])[4]
avgdaygap=0.76*summary(ad$gap[(as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour>=6 & (as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour<=16])[4]
avglngap=0.76*summary(ad$gap[(as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour>=22 | (as.POSIXlt(ad$datetime,format="%Y-%m-%d %H:%M:%S"))$hour<=05])[4]

#Calculate setpoints for generation
for (i in 1:nrow(ad)){
   if((as.POSIXlt(ad$datetime[i],format="%Y-%m-%d %H:%M:%S"))$hour>=17 & (as.POSIXlt(ad$datetime[i],format="%Y-%m-%d %H:%M:%S"))$hour<=21){
    ad$gensp[i]=avgevegap
    }
  if((as.POSIXlt(ad$datetime[i],format="%Y-%m-%d %H:%M:%S"))$hour>=6 & (as.POSIXlt(ad$datetime[i],format="%Y-%m-%d %H:%M:%S"))$hour<=16){
    ad$gensp[i]=avgdaygap
    }
  if((as.POSIXlt(ad$datetime[i],format="%Y-%m-%d %H:%M:%S"))$hour>=22 | (as.POSIXlt(ad$datetime[i],format="%Y-%m-%d %H:%M:%S"))$hour<=05){
    ad$gensp[i]=avglngap
  }
  #Update generation values based on the setpoints
  ad$newwindgen[i]=min(ad$windgen[i],ad$gensp[i])
  #Find total supply
  ad$totalsupply[i]=ad$newwindgen[i]+ad$blsp[i]
  #Find total delivered power
  ad$sputilized[i]=min(ad$demand[i],ad$totalsupply[i])
}

#Calculate curtailment
ad$spcurtailment=ad$totalsupply-ad$sputilized
#Calculate deficit
ad$spdeficit=ad$demand-ad$sputilized

#Print curtailment
sprintf("The total curtailment over the year, using setpoints, is %.2f kWh",sum(ad$spcurtailment))

#Calculate curtailed capacity factor
sum=0.0
for (i in 1:nrow(ad)){
sum=sum+max(ad$sputilized[i]-ad$blsp[i],0.0)
}
cf_c_sp=sum*100/(850*8760*7)
sprintf("Using setpoints for baseload and wind generation, curtailed CF is %.2f percent",cf_c_sp)
```
###**STORAGE CAPACITY EFFECTS ON CAPACITY FACTOR**
**5. Finally, perform the same calculations for your proposed storage capacity and set point values. How much additional energy is now recoverable via storage?**
```{r}
#Calculate state of  storage at each hour
ad$storage[1]=min(0.8*ad$spcurtailment[1],10000.0)
for (i in 2:nrow(ad)){
  if (ad$spdeficit[i]==0){
    ad$storage[i]=min((ad$storage[i-1]+0.8*ad$spcurtailment[i]),10000.0)
    }
  #With deficit
  else if (ad$spcurtailment[i]==0){
    ad$storage[i]=max(ad$storage[i-1]-((ad$spdeficit[i])/0.8),0.0)
    }
}

#Calculate net WE used in each hour
ad$we_used[1]=max(ad$sputilized[1]-ad$blsp[1],0.0)
#Find total wind energy supplied (including from storage) at each hour
for (i in 2:nrow(ad)){
  if (ad$spdeficit[i]==0){
    ad$we_used[i]=max(ad$sputilized[i]-ad$blsp[i],0.0)
    }
  else if (ad$spcurtailment[i]==0){
  ad$we_used[i]=max(ad$sputilized[i]-ad$blsp[i],0.0)+(ad$storage[i-1]-ad$storage[i])*0.8
  }
}


#Print previous useful WE
sprintf("Without storage, net useful wind energy is %.2f kWh",sum)

#Print new useful WE
sprintf("With storage, net useful wind energy is % .2fkWh",sum(ad$we_used))

#Print additional useful WE
sprintf("Therefore additional useful wind energy recoverable from storage is %.2f kWh",sum(ad$we_used)-sum)


#Print fraction of previous curtailment saved
sprintf("This is %.2f percent of the curtailment that would have occured without storage, %.2f kWh",(sum(ad$we_used)-sum)*100/sum(ad$spcurtailment),sum(ad$spcurtailment))


#Print fraction of previous unmet demand now met
sprintf("Therefore, %.2f kWh (%.2f percent) of non-baseload energy not met by the wind farm earlier, %.2f kWh, can be met by adding storage",sum(ad$we_used)-sum,(sum(ad$we_used)-sum)*100/sum(ad$spdeficit),sum(ad$spdeficit))

#Print new capacity factor
cf_c_bl_st_sp=sum(ad$we_used)*100/(850*8760*7)
sprintf("Capacity factor with storage is %.2f percent",cf_c_bl_st_sp)

plot3<-ggplot(ad, aes(x=ad$datetime)) 
plot3+geom_point(aes(y=ad$demand),colour="blue",size=1)+geom_point(aes(y=ad$blsp),colour="black",size=1)+geom_point(aes(y=ad$blsp+ad$gensp),colour="red",size=1)+ylab("Energy (kWh)")+xlab("Time of the Year")+ggtitle("Set Points against Demand\n Baseload (black), Generation (red) and Demand (blue)")

plot4<-ggplot(ad, aes(x=ad$datetime))
plot4+geom_line(aes(y=ad$demand),colour="blue",size=1)+geom_line(aes(y=ad$blsp),colour="black",size=1)+geom_line(aes(y=ad$totalsupply),colour="red",size=1)+ylab("Energy (kWh)")+xlab("Time of the Year")+ggtitle("Supply and Demand\n Baseload (black), Total supply (red) and Demand (blue)")
```

